%==============================================================================
% Sjabloon onderzoeksvoorstel bachproef
%==============================================================================
% Gebaseerd op document class `hogent-article'
% zie <https://github.com/HoGentTIN/latex-hogent-article>

% Voor een voorstel in het Engels: voeg de documentclass-optie [english] toe.
% Let op: kan enkel na toestemming van de bachelorproefcoördinator!
\documentclass[english]{hogent-article}

% Invoegen bibliografiebestand
\addbibresource{voorstel.bib}

% Informatie over de opleiding, het vak en soort opdracht
% \studyprogramme{Professionele bachelor toegepaste informatica}
% \course{Bachelorproef}
% \assignmenttype{Onderzoeksvoorstel}
% Voor een voorstel in het Engels, haal de volgende 3 regels uit commentaar
\studyprogramme{Bachelor of applied information technology}
\course{Bachelor thesis}
\assignmenttype{Research proposal}

\academicyear{2023-2024}

\title{Server-side Programming Languages for Scripting Customizations and Extensions to IAM Systems: A Comparative Analysis and Proof of Concept}

\author{Aiko Vandeputte}
\email{aiko.vandeputte@student.hogent.be}

\supervisor[Co-promotor]{Y. De Backer (TrustBuilder, \href{mailto:ydb@trustbuilder.com}{ydb@trustbuilder.com})}

% Binnen welke specialisatierichting uit 3TI situeert dit onderzoek zich?
% Kies uit deze lijst:
%
% - Mobile \& Enterprise development
% - AI \& Data Engineering
% - Functional \& Business Analysis
% - System \& Network Administrator
% - Mainframe Expert
% - Als het onderzoek niet past binnen een van deze domeinen specifieer je deze
%   zelf
%
\specialisation{Mobile \& Enterprise development}
\keywords{Identity and Access Management, Customization, Scripting}

\begin{document}

\begin{abstract}
  Identity and Access Management (IAM) systems need tight integration with user-facing applications and other systems. To achieve this, IAM services allow writing scripts that can customize and extend their systems. TrustBuilder, a European IAM company, wants to overhaul their customization system, and needs to know what language(s) they should allow these scripts to be written in. This research will investigate which languages would be the best choice for scripting extensions and customizations for IAM platforms, and aims to arrive at a singular primary language choice. A comparative study will be conducted to evaluate the suitability of different languages for IAM tasks, and a proof of concept will be built which will allow running scripts that can handle those tasks in the ultimately chosen language(s). Being originally designed for the web, as well as being the most widely-used language, JavaScript is expected to end up as the ideal language. However, as more exhaustive research is done, important criteria such as security and reliability implications may result in a different candidate being chosen as the best fit. A final language choice, with proven suitability for IAM tasks, will allow TrustBuilder to rework their customization tooling with certainty.
\end{abstract}

\tableofcontents

% De hoofdtekst van het voorstel zit in een apart bestand, zodat het makkelijk
% kan opgenomen worden in de bijlagen van de bachelorproef zelf.
\input{voorstel-inhoud}

\printbibliography[heading=bibintoc]

\end{document}