%---------- Inleiding ---------------------------------------------------------

\section{Introduction}%
\label{sec:introduction}

This research proposal is written based on a business case of TrustBuilder, a European company providing Identity and Access Management (IAM) solutions for organizations across Europe, with extensive customization capabilities to satisfy client-specific needs.

Their product, also named TrustBuilder, is platform that aims to be able to handle any IAM need, such as authentication, session management, authorization, and user management. Further requirements can take many forms, and the TrustBuilder platform needs to be able to tightly integrate with each client’s user-facing applications and other services. As these needs may differ from application to application, not everything can be a native part of the IAM system, and customization capabilities are a must. In order to satisfy these needs, their clients need to be able to to customize and extend their platform using custom code.

To tackle this problem, TrustBuilder allows for the creation of “workflows”. These workflows consist of simple building blocks such as “script blocks”, “adapters” and “services”, using arrow connectors and “condition” blocks between them for control flow. Script blocks can run custom JavaScript functions, allowing workflows to be used to implement virtually any functionality missing from the base product, as well as integrate with the base product directly.

TrustBuilder wants to overhaul their workflow platform to improve the developer experience and maintainability of workflows. Therefore, they want to know what language(s) their workflow platform should be built around.

This thesis will investigate which languages would be the best choice for scripting extensions and customizations for IAM-platforms. A primary language choice based on the performed research and a proof of concept showcasing the language’s suitability for IAM-related tasks will allow TrustBuilder to start the overhaul of workflows with certainty.

%---------- Stand van zaken ---------------------------------------------------

\section{State of the art}%
\label{sec:state-of-the-art}

Hier beschrijf je de \emph{state-of-the-art} rondom je gekozen onderzoeksdomein, d.w.z.\ een inleidende, doorlopende tekst over het onderzoeksdomein van je bachelorproef. Je steunt daarbij heel sterk op de professionele \emph{vakliteratuur}, en niet zozeer op populariserende teksten voor een breed publiek. Wat is de huidige stand van zaken in dit domein, en wat zijn nog eventuele open vragen (die misschien de aanleiding waren tot je onderzoeksvraag!)?

Je mag de titel van deze sectie ook aanpassen (literatuurstudie, stand van zaken, enz.). Zijn er al gelijkaardige onderzoeken gevoerd? Wat concluderen ze? Wat is het verschil met jouw onderzoek?

Verwijs bij elke introductie van een term of bewering over het domein naar de vakliteratuur, bijvoorbeeld~\autocite{Hykes2013}! Denk zeker goed na welke werken je refereert en waarom.

Draag zorg voor correcte literatuurverwijzingen! Een bronvermelding hoort thuis \emph{binnen} de zin waar je je op die bron baseert, dus niet er buiten! Maak meteen een verwijzing als je gebruik maakt van een bron. Doe dit dus \emph{niet} aan het einde van een lange paragraaf. Baseer nooit teveel aansluitende tekst op eenzelfde bron.

Als je informatie over bronnen verzamelt in JabRef, zorg er dan voor dat alle nodige info aanwezig is om de bron terug te vinden (zoals uitvoerig besproken in de lessen Research Methods).

% Voor literatuurverwijzingen zijn er twee belangrijke commando's:
% \autocite{KEY} => (Auteur, jaartal) Gebruik dit als de naam van de auteur
%   geen onderdeel is van de zin.
% \textcite{KEY} => Auteur (jaartal)  Gebruik dit als de auteursnaam wel een
%   functie heeft in de zin (bv. ``Uit onderzoek door Doll & Hill (1954) bleek
%   ...'')

Je mag deze sectie nog verder onderverdelen in subsecties als dit de structuur van de tekst kan verduidelijken.

%---------- Methodologie ------------------------------------------------------
\section{Methodology}%
\label{sec:methodology}

The research will be divided into 6 phases, the first of which is a literature review. Literature research will be performed to explore existing approaches, find expert opinions, and gain a deeper understanding of the state of the art. The result of this literature search can be found in section 2 of this research proposal.

Phase 2 consists of creating a list of minimal requirements for a suitable language choice. Possible types of requirements could include ease of use, security implications, performance, and suitability for IAM-related tasks. A concrete list of measurable requirements will be assembled in consultation with thesis stakeholders and employees at TrustBuilder. The MoSCoW method will be used to order these requirements by importance, and each requirement will be marked as functional or non-functional.

Phase 3 is to create a long list of options, checked against the previously listed requirements. The literature review from phase 1 will be used to build an exhaustive list of suitable language choices. Popular existing options, as well as possibly lesser-known choices, will be considered.

Phase 4 aims to create a short list of the three most promising options, using a requirements summary table.

In phase 5, a proof of concept will be created to validate the suitability of each language in the short list. The proof of concept will consist of a simple HTTP server for each chosen language, which will allow running code that handles HTTP requests in the specified language. Three scripts will be written for each language, and will handle three distinct needs. The exact functionality of these three types of scripts will be decided based on TrustBuilder's typical customization requirements. This proof of concept will help further solidify and compare each language's suitability for customizing and extending IAM systems, and will provide deeper insight in their tooling and ease of use for these purposes.

Phase 6 will consist of drawing conclusions out of the research and propose next steps. The results and findings from the previous phases will be analyzed, and limitations and areas for future research will be identified.

Throughout this entire process, a thesis will be written that documents the research, findings, and proposed solutions.


%---------- Verwachte resultaten ----------------------------------------------
\section{Expected result, conclusion}%
\label{sec:expected_results}

Based on the findings in the literature review, JavaScript/TypeScript is expected to be the ideal language for writing extensions for IAM systems. It is currently the most popular programming language, and is already widely used for the extension of platforms, both in the general web development space as for IAM-related tasks. Being the most popular language, it has a rich package ecosystem, is actively being developed, and has a range of possible runtime environments. Having been specifically created for the web, with new standardized web APIs still being added, it can easily handle OAuth and SAML requirements, as both technologies were designed on top of existing web standards. JavaScript is already an excellent language choice for these purposes, and its continued innovation cements its place for years to come.

As many languages can be compiled to JavaScript or WASM now, an extension system built on top of a JavaScript and WASM engine would also enable the usage of other languages, as technologies and requirements continue to evolve. Python remains a popular and easy-to-use language for scripting, and Rust is frequently used to safely write critical code. A system that enables this level of versatility would provide a future-proof platform for mission-critical customization scripts.

